---
title: "Intro to R Assignment"
author: "Annie Britton"
date: "1/23/2023"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Load dplyr.**
```{r message=FALSE, warning=FALSE}
library(dplyr)
```

**Set working directory to activity folder.**
```{r}
setwd("C:/Users/annie/Documents/AEEN/M01/Activity1")
```

**Import the stream data set and save as stream1.**
```{r}
stream1 <- read.csv(file="assignment1_data.csv", header=TRUE)
```

**Group the data according to stream number and calculate the mean of fish density (number_fish column) for each stream.**
```{r}
stream1 %>% 
  group_by(stream_number) %>% 
  summarise(mean_fish_density = mean(number_fish))
```

**Group the data according to stream type and calculate the mean of fish density for each stream type.**
```{r}
stream1 %>% 
  group_by(stream_type) %>% 
  summarise(mean_fish_density = mean(number_fish))
```

**Create your first vector using the stream number and call that vector ‘a’.**
```{r}
a <- stream1 %>% 
  group_by(stream_number) %>% 
  summarise(mean_fish_density = mean(number_fish)) %>% 
  pull(stream_number)
```

**Create your second vector using the mean of fish density you calculated for each stream and call that vector ‘b'.**
```{r}
b <- stream1 %>% 
  group_by(stream_number) %>% 
  summarise(mean_fish_density = mean(number_fish)) %>% 
  pull(mean_fish_density)
```

**Make a data frame called ‘stream2’ using these two vectors you just created.**
```{r}
stream2 <- tibble(stream_number = a, mean_fish_density = b)
stream2
```

**Make a graph using this new data with the help of plot() function.**
```{r}
plot(stream2, main="Fish Density by Stream",
     xlab="Stream Number",
     ylab="Fish Density Measurements (fish/m2)")
```
